---
title: "Knapsack"
author: Olof Josefsson
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Knapsack}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(knapsack)
```

# 1.1.2
```{r}
library(knapsack)
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
data.frame(
w=sample(1:4000, size = n, replace = TRUE),
v=runif(n = n, 0, 10000)
)

system.time(brute_force_knapsack(x = knapsack_objects[1:16,], W = 3500))
```
# 1.1.3
```{r}
library(knapsack)
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
data.frame(
w=sample(1:4000, size = n, replace = TRUE),
v=runif(n = n, 0, 10000)
)

system.time(knapsack_dynamic(x = knapsack_objects[1:500,], W = 3500))
```
# 1.1.4
```{r}
library(knapsack)
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 1000000
knapsack_objects <-
data.frame(
w=sample(1:4000, size = n, replace = TRUE),
v=runif(n = n, 0, 10000)
)

system.time(greedy_knapsack(x = knapsack_objects[1:1000000,], W = 3500))
```

# 1.1.6 Profiling
## Brute force
For the brute force algorithm, the part that was registered as something that took
memory and time while profiling was the following line of code:
`weight <- sum(w[m==1])`.

Replacing this with `weight <- sum(unlist(w[m==1], use.names=FALSE))` made the code run faster and no trace of time or memory consumption was thereafter shown.

## Dynamic 
For the dynamic knapsack implementation the thing that took allocated memory and 
time was `m[i, j] = max(m[i-1, j], m[i-1,j-w[i-1]]+v[i-1])`. However no solution
to a better way of implementing this piece of the function was found.

## Greedy
For the profiling of the greedy function a bottleneck was identified in the following line of code:
`x<-x[rev(order(x$ratio)),]`

Below is how I tried to fix this using microbenchmark package. **It takes some time to run**

```{r}
library(knapsack)
library(microbenchmark)
# Set up data
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 1000000
knapsack_objects <-
  data.frame(
  w=sample(1:4000, size = n, replace = TRUE),
  v=runif(n = n, 0, 10000)
)
# Add df column to knapsack_objects.
knapsack_objects["ratio"] = knapsack_objects$v/knapsack_objects$w

##### Functions for testing #####
way1 <- function(x = knapsack_objects[1:1000000,]){
  x<-x[rev(order(x$ratio)),]
  return(x)}
way2 <- function(x = knapsack_objects[1:1000000,]){
  x<- x[order(x$ratio),]
  x<- x[nrow(x):1,]
  return(x)}
way3 <- function(x = knapsack_objects[1:1000000,]){
  x<- x[order(-x$ratio), ]
  return(x)}

a <- way1()
b <- way2()
c <- way3()

##### Make sure output is the same #####
all.equal(a, b)
all.equal(a, c)

##### Output of benchmarking #####
microbenchmark::microbenchmark(way1(),way2(), way3(), times = 100L)
```

Testing this way found that `way1()` was the most stable way of doing it, although `way3()`
sometimes had the shortest run time. Therefor this bottleneck was not corrected in any way,
but rather left as it was before the profiling/benchmarking.

## General comment after profiling
The overall run time and allocated memory of the different functions tended to be rather
fast -> very few identifications of potential bottlenecks. Therefor the performance gain in
this particular lab was minimal. Nonetheless it seems like a great tool for writing faster code (especially the `microbencmark` package).

# 1.1.8 
For the parallel part of the brute force algorithm, `parLapply`
was implemented. This was chosen as it was supported by windows OS. A performance gain 
was not gained via the parallel implementation, as generating the needed cluster takes a lot of time.
